---
title: "R Homework 2"
author: "Yichuan Zhang 47-216786"
date: "10/12/2021"
output: 
  pdf_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    toc: yes
    toc_depth: 4
---

# compute percentage college graduate by US state

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(gapminder)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(Hmisc)
library(wesanderson)
library(ggridges)
library(griffen)
```


dplyr的基本用法:1）filter 2)select 3)mutate 4)arrange 5)summarise 6)group_by

```{r}
options(width = 100)
# show what variables we have
names(cps)
```

## select variables

At here we use **select** method.
```{r}
options(width = 100)
# we need to select the variables we need
new_df <- cps %>% select(state, education_category)
new_df
```

## groupby and summarise

At here we use **group_by** and **summarise**

```{r}
options(width = 100)

# group by the state and educatio_category and count all the categories

#count_df <- new_df %>% group_by(state, education_category) %>% summarise(n = n()).count()

count_df <- count(new_df %>% group_by(state, education_category))

count_df
```

# Loop to get the graduation rate for each state

At here we need a loop the run for each state, and **mutate** will be used at here

```{r}
group_df <- count_df %>% filter(state == "Ohio")
total_number_df <- group_df %>% summarise(total_num = sum(n))
total_number <- as.integer(total_number_df["total_num"] %>% summarise(total_student = sum(total_num)))
graduation_rate_df <- count_df %>% filter(state == "Ohio") %>% mutate(graduation_rate=n/total_number)
graduation_rate_df
state_graduation_rate <- graduation_rate_df["graduation_rate"][[1]][1]
```

```{r}
options(width = 100)
# get the unique states
unique_state <- unique(new_df["state"])[[1]]
# create an empty list
desired_length <- 1
graduation_rate <- rep(NA, desired_length)

for (i in unique_state) {
    group_df <- count_df %>% filter(state == i)
    total_number_df <- group_df %>% summarise(total_num = sum(n))
    total_number <- as.integer(total_number_df["total_num"] %>% summarise(total_student = sum(total_num)))
    graduation_rate_df <- count_df %>% filter(state == i) %>% mutate(graduation_rate=n/total_number)
    state_graduation_rate <- graduation_rate_df["graduation_rate"][[1]][1]
    graduation_rate <- c(graduation_rate, state_graduation_rate)
}

# remove the first na value in the list
graduation_rate <- graduation_rate[-1]
graduation_rate
```

# create the dataframe for drawing the plot

```{r}
options(width = 100)
# make a new dataframe
final_df <- data.frame(unique_state, graduation_rate)
# order the dataframe
final_df <- final_df[order(graduation_rate,
 decreasing = FALSE),]
# rename the index
rownames(final_df) <- 1 : length(rownames(final_df))
# avoid the ggplot sort the geom_point automaticlly
final_df$unique_state <- factor(final_df$unique_state,
    levels = final_df$unique_state)
final_df
```

# Draw the graph
```{r}

p <- ggplot(data=final_df,
 mapping = aes(x=unique_state, y=graduation_rate)) +
  geom_point() + coord_flip() + labs(y = "College Graduation Rate", x = "")
p
```

